```{r knitsetup, echo=FALSE, results='hide', warning=FALSE, message=FALSE, cache=FALSE}
opts_knit$set(base.dir='./', fig.path='', out.format='md')
opts_chunk$set(prompt=FALSE, comment='', fig.align = 'center', results='markup')
```

# Extract element

In some cases, one might need to extract an element from an object. For example, extract `r.squared` from the summary of a linear model, extract the test statistics from a S4 object.

The syntax `x %>>% (y)` indicates to extract the element named `y` from object `x` where `y` must be a valid symbol name and `x` can be a vector, list, environment or anything else that supports `[[]]`. It also works with S4 object and able to extract a slot from it in exactly the same way.

```{r}
mtcars %>>%
  (lm(mpg ~ wt + cyl, data = .)) %>>%
  summary %>>%
  (r.squared)
```

The syntax is not ambiguous with other features that used `()` because evaluating a symbol simply makes no sense.

As we know that `[[]]` are defined for vector, list, and environment, the following simple examples work.

```{r}
list(a=1,b=2) %>>% (a) # list(a=1,b=2)[["a"]]
c(a=1, b=2) %>>% (a) # c(a=1,b=2)[["a"]]
```

For environment,

```{r}
env <- new.env()
env$a <- 1
env$b <- 2
# do more things here ...
env %>>% (a) # env$a
```

For S4 object, the syntax works too. We use `adfTest()` in library `fUnitRoots` to perform an ADF test to test the cointegration relationship between two vectors.

Suppose the data generating process is

```{r extract-s4, fig.height = 4}
set.seed(0)
x <- cumsum(rnorm(150))
y <- 2*x + rnorm(150)*0.5
matplot(cbind(x,y),type = "l")
```

Now given the data, we can perform an ADF test on the residues of linear model `lm(y ~ x + 0)`.

```{r}
lm(y ~ x + 0) %>>%
  residuals %>>%
  fUnitRoots::adfTest(type = "nc") %>>% # S4 object
  (~ testobj) %>>% # save test object to environment
  (test) %>>% # testobj@test
  (statistic) # testobj@test$statistic
```

Note that the result of `fUnitRoots::adfTest()` is a S4 object which is assigned to `testobj`, and the result of `(test)` is no longer a S4 object but an ordinary list. Therefore, there's no different in the syntax of element extraction between S4 object and other objects that support `[[]]`.

Since the syntax only accepts symbol name, to evaluate an expression with the piped object being the frame if it is a list or environment, using `with()` or `within()` can be helpful.

```{r}
list(a = 1, b = 2) %>>%
  with(a+2*b)
```

But this method does not work for vector and S4 object.
