```{r knitsetup, echo=FALSE, results='hide', warning=FALSE, message=FALSE, cache=FALSE}
opts_knit$set(base.dir='./', fig.path='', out.format='md')
opts_chunk$set(prompt=FALSE, comment='#', fig.align = 'center', results='markup')
library(pipeR)
```

# Pipe to dot

Not all functions are pipe-friendly in every case: You may find some functions do not take your data produced by a pipeline as the first argument. In this case, you can enclose your expression by `{}` or `()` so that `%>>%` will use `.` to represent the value on the left and evaluate the enclosed expression.

For the linear model example, one can rewrite `lm(formula = )` to

```{r}
mtcars %>>%
  { lm(mpg ~ cyl + wt, data = .) }
```

or 

```{r}
mtcars %>>%
  ( lm(mpg ~ cyl + wt, data = .) )
```

The difference between `{}` and `()` used above is

1. `{}` accepts more than one expressions within the braces and its value is determined by the last one; but `()` accepts only one expression.
2. `{}` has only one feature: pipe to `.` in the enclosed expression while `()` has more features (we will cover them soon).

To demonstrate the difference, we run the following examples.

```{r}
mtcars %>>% {
  model <- lm(mpg ~ wt + cyl, data = .)
  summ <- summary(model)
  summ[c("r.squared","adj.r.squared")]
}
```

It shows that `{}` holds a code block whose value is determined by the last expression or interrupted by `return()`.

```{r}
mtcars %>>% 
  lm(formula = mpg ~ wt + cyl) %>>%
  summary %>>% {
    if(.$r.squared >= 0.8) {
      return("Model A")
    }
    cat("Model B should be considered.\n")
    "Model B"
  }
```

One thing to notice is that `{}` is more flexible than previously demonstrated. It also allows using `%>>%` within the braces as well as causing side effect such as plotting graphics.

```{r multiline, fig.height=4}
mtcars %>>% {
  par(mfrow=c(1,2))
  .$mpg %>>% plot(col = "red", main="mtcars (mpg)")
  .$mpg %>>% hist(main = "distribution")
}
```

Another thing is that the meaning of nested dots are always determined by the working operator for that dot symbol. For example, the above code can be refined to

```{r nested_dots, fig.height=4}
mtcars %>>% {
  par(mfrow=c(1,2))
  .$mpg %>>% plot(col = "red", main=sprintf("mtcars (mpg: %d)",length(.)))
  .$mpg %>>% hist(main = "distribution")
}
```

It should be obvious that `.` below `par()` belong to the first `%>>%` that works with `mtcars` while `.` in `length()` belong to the operator that works with `.$mpg` so that it can correctly show the length of `mpg` (`r length(mtcars$mpg)`) rather than that of `mtcars` (`r length(mtcars)`).

## Creating closure

The expression will be evaluated with speically defined `.` if it is enclosed by `{}` or `()`. If an anonymous function, that is, a function definition without a name, is enclosed, the pipeline will result in a closure, a function returned by a function (`%>>%`) which can be used more times.

In the following example, we create a closure that takes one argument: the kernel method with which to perform the nonparametric density estimation of resampled `mtcars$mpg`.

```{r}
density_plot <- mtcars$mpg %>>%
  sample(size = 10000, replace = TRUE) %>>%
  (function(kernel) {
    . %>>%
      density(kernel = kernel) %>>%
      plot(main = sprintf("%s kernel", kernel))
  })
```

We just create a closure called `density_plot`. Then we can call it with different kernels to see the differece between the density plots of each other.

```{r create-closure, fig.height=4, fig.width=10}
par(mfrow=c(1,3))
density_plot("gaussian")
density_plot("rectangular")
density_plot("triangular")
```

However, if the enclosed function is called directly in the pipeline, `%>>%` will pipe the left value to its first argument because here the only difference between first-argument piping examples and this example is the function no longer has a name but created inline.

```{r}
1:10 %>>% (function(x,pow) x^pow)(2)
```

A more complicated example is 

```{r}
mtcars %>>%
  lm(formula = mpg ~ wt + cyl) %>>%
  (function(model, warn_level) {
    if(summary(model)$r.squared < warn_level)
      warning("r.squared is too low", call. = FALSE)
    model
  })(0.9) %>>%
  coef
```

where an anonymous function is defined inline and `%>>%` following `lm()` pipes the linear model to the first argument of the anonymous function.
