```{r knitsetup, echo=FALSE, results='hide', warning=FALSE, message=FALSE, cache=FALSE}
opts_knit$set(base.dir='./', fig.path='', out.format='md')
opts_chunk$set(prompt=FALSE, comment='', fig.align = 'center', results='markup')
```

# Pipe to first argument

Many R functions are pipe-friendly: they take some data by the first argument and transform it in a certain way. This arrangement allows operations to be streamlined by pipes, that is, one data source can be put to the first argument of a function, get transformed, and put to the first argument of the next function. In this way, a chain of commands are connected, and it is called a pipeline.

Here is an example of reorganizing code in pipeline written with elementary functions.

Suppose the original code is

```{r}
summary(sample(diff(log(rnorm(100,mean = 10))),
  size = 10000,replace = TRUE))
```

Note that `rnorm()`, `log()`, `diff()`, `sample()`, and `summary()` all take the data as the first argument. We can use `%>>%` to rewrite the code so that the process of data transformation is straightforward.

```{r}
library(pipeR)
set.seed(123)
rnorm(100, mean = 10) %>>%
  log %>>%
  diff %>>%
  sample(size = 10000, replace = TRUE) %>>%
  summary
```

The syntax of first argument piping is that, on the right-hand side of `%>>%`, whenever a function name or call is supplied, the left-hand side value will always be put to the first unnamed argument to that function.

| Syntax | Evaluate as |
|--------|-------------|
|`x %>>% f`| `f(x)` |
|`x %>>% f(...)`| `f(x,...)` |

Although you can write everything in one line but that is probably not very elegant. It is better to be generous to trade readability with the number of lines. 

Note that, at each line, whenever we want to continue building the pipeline with the previous result, we end that line with `%>>%`. If one line does not end up with `%>>%`, the pipeline ends.

Some more examples with graphics functions:

```r
mtcars$mpg %>>%
  plot
```

```r
mtcars$mpg %>>%
  plot(col="red")
```

Sometimes the value on the left is needed at multiple places. In this case you can use `.` to represent it anywhere in the function call.

```r
mtcars$mpg %>>%
  plot(col="red", main=sprintf("number of points: %d",length(.)))
```

There are situations where one calls a function in a namespace with `::`. In this case, the call must end up with parentheses with or without parameters..

```r
mtcars$mpg %>>%
  stats::median()
  
mtcars$mpg %>>%
  graphics::plot(col = "red")
```

Notice that `%>>%` not only works between function calls, but also can be nested in function calls. For example,

```{r}
mtcars %>>%
  subset(mpg <= quantile(mpg,0.95), c(mpg, wt)) %>>%
  summary
```

can be written like 

```{r}
mtcars %>>%
  subset(mpg <= mpg %>>% quantile(0.95), c(mpg, wt)) %>>%
  summary
```

In some other cases, the function is not very friendly to pipeline operation, that is, it does not take the data you transform through a pipeline as the first argument. One example is the linear model function `lm()`. This function take `formula` first and then `data`.

If you directly run

```{r}
mtcars %>>%
  lm(mpg ~ cyl + wt)
```

it will fail because `%>>%` is evaluating `lm(mtcars, mpg ~ cyl + wt)` which does not fulfil the expectation of the function. There are two ways to build pipeline with such kind of functions.

First, use named parameter to specify the formula.

```{r}
mtcars %>>%
  lm(formula = mpg ~ cyl + wt)
```

This works because it is actually evaluated as 

```r
lm(mtcars, formula = mpg ~ cyl + wt)
```

and R's argument matching program decides that since the first argument in `lm()`'s definition `formula` is specified, the unnamed argument `mtcars` is regarded as specifying the second argument `data`, which is exactly what we want. Therefore, it works fine here.

However, this trick only makes it easy for some functions but not all. Suppose a function that takes `data` as the third or fourth argument. In this case, you would have to explictly specify all previous arguments by name. If `data` argument follows `...`, the trick would not work at all.

Dot piping is designed for more flexible pipeline construction. It allows you to use `.` to represent the left-hand side value and put it anywhere you want in the next expression. The next page demonstrates its syntax and when it might be useful.
