```{r knitsetup, echo=FALSE, results='hide', warning=FALSE, message=FALSE, cache=FALSE}
opts_knit$set(base.dir='./', fig.path='', out.format='md')
opts_chunk$set(prompt=FALSE, comment='#', fig.align = 'center', results='markup')
library(pipeR)
```

# Dot function

As we have mentioned earlier, not all functions are pipe-friendly. While `%>>%` accepts `{}` and `()` on the right side, `Pipe` object provides `.()` to perform exactly the same set of features with `x %>>% (...)` which is powerful enough to meet a wide variety of demands in piping.

## Pipe to dot

If you supply an ordinary function call in `.()` it will evaluate it with `.` being the piped value.

```{r}
Pipe(mtcars)$
  .(lm(mpg ~ cyl + wt, data = .))
```

## Pipe by formula

To avoid ambiguity, one can write a formula in `.()` to define the symbol representing `mtcars`.

```{r}
Pipe(mtcars)$
  .(df ~ lm(mpg ~ cyl + wt, data = df))
```

## Pipe for side effect

The syntax of one-sided formula is also interpreted as side effect in `.()`.

```{r}
Pipe(mtcars)$
  subset(mpg >= quantile(mpg, 0.05) & mpg <= quantile(mpg, 0.95),
    c(mpg, cyl, wt))$
  .(~ cat("rows:",nrow(.),"\n"))$
  lm(formula = mpg ~ .)$
  summary()$
  coef()
```

The question mark works too.

```{r}
Pipe(mtcars)$
  subset(vs == 1, c(mpg, cyl, wt))$
  .(? nrow(.))$
  .(? data ~ ncol(data))$
  summary()
```

The debugging function `browser()` works exactly the same.

```r
mtcars %>>% 
  subset(vs == 1, c(mpg, cyl, wt)) %>>%
  (~ browser()) %>>%
  lm(formula = mpg ~ cyl + wt) %>>%
  summary
```

## Pipe with assignment

As we have previously mentioned that assignment may be the most important side effect so it deserves a syntax. One-sided formula with a symbol on the right indicates the input value should be assigned to that symbol. The syntax works in `.()` too.

```{r}
pmodel <- Pipe(mtcars)$
  subset(select = c(mpg, wt, cyl))$
  .(~ sub_mtcars)$ # assign subsetted mtcars to sub_mtcars
  lm(formula = mpg ~ wt + cyl)

summary(sub_mtcars)
```

Note that `=` in function argument will be interpreted as argument selection, therefore `=` cannot be used to assign value in `.()` function.

The way to perform side-effect assignment is to use `<-` or `->` introduced by pipeR v0.5 with `~`, like the following:

```{r}
Pipe(mtcars)$
  lm(formula = mpg ~ wt + cyl)$
  .(~ summary(.) -> summ)$
  coef()

summ$r.squared
```

If `~` in is removed, the assignment will be done and the value of `summary()` will go on piping.

```{r}
Pipe(mtcars)$
  lm(formula = mpg ~ wt + cyl)$
  .(summary(.) -> summ)$
  coef()

summ$r.squared
```

Note that the output of two `Pipe`s are different. The `Pipe` with `~ summary(.)` results in a numeric vector of the coeffecients of the linear model because the summary is only a side effect while the `Pipe` without `~` results in a comprehensive matrix showing the summary of the linear coefficients because `coef()` works on the summary.

## Extract element

Element extraction also works in `.()` if a symbol name is supplied.

```{r}
Pipe(mtcars)$
  lm(formula = mpg ~ wt + cyl)$
  summary()$
  .(r.squared)
```

## Dot function as closure

Note that `.()` is only defined within a `Pipe` object and is not an exported function in pipeR, you cannot directly call `.()` outside a `Pipe()`. 

In fact, if you are familiar with functional programming, it is called a *closure*, which is basically defined as a function returned by a function. However, you can save `.()` in a `Pipe` object to a symbol as a shortcut to call it elsewhere.

```{r}
model <- Pipe(mtcars)$
  lm(formula = mpg ~ wt + cyl)$
  .

model()
model(coefficients)
model()$summary()$coef()
```
