```{r knitsetup, echo=FALSE, results='hide', warning=FALSE, message=FALSE, cache=FALSE}
opts_knit$set(base.dir='./', fig.path='', out.format='md')
opts_chunk$set(prompt=FALSE, comment='#', fig.align = 'center', results='markup')
```

# Pipe to first argument

`Pipe()` creates a Pipe object that supports light-weight chaining with `$`. Basically, it is like a box containing a value and for this box `$` is defined to perform first-argument piping.

The example in the page that introduces the first-argument piping feature of `%>>%` can be translated using `Pipe()` and `$`:

```{r}
library(pipeR)
set.seed(123)
Pipe(rnorm(100, mean = 10))$
  log()$
  diff()$
  sample(size = 10000, replace = TRUE)$
  summary()
```

You may notice that in the beginning of the pipeline, the numeric vector generated by `rnorm()` is converted to a `Pipe` object by `Pipe()`. Since `$` for `Pipe` object is defined for first-argument piping, the name following `$` will be interpreted to a function name, and then `()` will execute that function with the value in `Pipe` put to the first argument, and put the result into the next `Pipe` object. In this way, the pipeline continues chaining.

In other words, `Pipe(x)$f()` works like `Pipe(f(x))`, and `Pipe(x)$f(a)$g(b)` works like `Pipe(g(f(x,a),b))`, which can be further grown and become more nested if written without `$` chaining. 

Also note that the output does not look exactly the same with that produced by `%>>%` but with a header added on top of the summary table. Recall the version of code using the operator:

```{r}
set.seed(123)
rnorm(100, mean = 10) %>>%
  log %>>%
  diff %>>%
  sample(size = 10000, replace = TRUE) %>>%
  summary
```

In fact, the result produced by `Pipe()` is not really the summary table but a box (`Pipe` object) that contains it, and therefore the object can continue piping with `$`. `Pipe` object implements several generic functions to make it easier to inspect and manipulate the value in the box. To extract the inner value, call `$value`, or simply `[]` as shortcut.

```{r}
set.seed(123)
Pipe(rnorm(100, mean = 10))$
  log()$
  diff()$
  sample(size = 10000, replace = TRUE)$
  summary()$
  value
```

With `$value` in the end of the pipeline, `value` is extracted from the Pipe object, and this object is no longer a `Pipe` so that `$` won't pipe any more.

Here is another example that runs a linear regression with `mtcars`.

```{r}
Pipe(mtcars)$
  lm(formula = mpg ~ wt + cyl)
```

Use `[]` to extract the `lm` object inside the resulted `Pipe` object.

```{r}
Pipe(mtcars)$
  lm(formula = mpg ~ wt + cyl) []
```

The Pipe header no longer exists, which indicates that the value is extracted.

Just like `%>>%`, in the function call after `$` also supports `.` to represent the input value. For example,

```{r pipe-dot, fig.height=4}
Pipe(mtcars$mpg)$
  plot(col = "red", main = sprintf("mpg (%d obs.)", length(.)))
```

You may notice that the previous `plot()` only produces graphics but the `NULL` value it returns are not explicitly printed. `Pipe` by design mutes `NULL` value as being printed. However, not all graphics functions return `NULL`. `hist()` is one example.

```{r pipe-hist, fig.height=4}
Pipe(mtcars$mpg)$
  hist(main = "distribution of mpg")
```

The output is no longer `NULL` but a new `Pipe` object consisting of a `histogram` object with a few elements indicating its properties. 

All `Pipe` objects are printed with, by default, a header like `$value: class`. If you find it annoying, you can turn off the header by setting the option `Pipe.header` to `FALSE` with

```r
options(Pipe.header = FALSE)
```

This is **NOT** recommended because `Pipe` object and ordinary objects are essentially different. For better distinction, we suggest that you give  all `Pipe` objects name that start with `p` and do not turn off this option.

The following example demonstrate a recommended use of `Pipe` object in multiple ways.

```{r}
pmtcars <- Pipe(mtcars)$
  subset(mpg >= quantile(mpg, 0.05) & mpg <= quantile(mpg, 0.95))

pmtcars$
  lm(formula = mpg ~ wt + cyl)$
  coef()

pmtcars$
  lm(formula = mpg ~ wt + cyl + qsec)$
  summary()$
  coef()
```

Note that we create a `Pipe` object from `mtcars` and filters it by lower and upper quantile. The result is still a `Pipe` object so that we can pipe with it further until we use `$value` or `[]` to extract its value.

## Creating partial function

```{r partial-function, fig.height=4}
density_plot <- Pipe(mtcars$mpg)$
  sample(size = 10000, replace = TRUE)$
  density(kernel = "gaussian")$
  plot

par(mfrow=c(1,2))
density_plot(col = "blue", main = "blue points")
density_plot(col = "gray", type = "o", main = "gray circles")
```

