```{r knitsetup, echo=FALSE, results='hide', warning=FALSE, message=FALSE, cache=FALSE}
opts_knit$set(base.dir='./', fig.path='', out.format='md')
opts_chunk$set(prompt=FALSE, comment='#', fig.align = 'center', results='markup')
```

# rlist

[rlist](http://renkun.me/rlist) is a set of tools for working with list objects. Its goal is to make it easier to work with lists by providing a wide range of functions on non-tabular data stored in them.

Its design is very much like dplyr but rlist is focused at dealing with non-tabular data stored in JSON or YAML format which can be easily represented by list objects in R.

In this example, we will get to know more about [Hadley Wickham](https://github.com/hadley)'s contribution to R community from JSON data retrieved with [GitHub API](https://api.github.com) using rlist and pipeR.

```{r}
library(pipeR)
library(rlist) # devtools::install_github("rlist","renkun-ken")
url <- "https://api.github.com/users/hadley/repos?per_page=100&page=%d"
repos <- 1:2 %>>%
  list.map(list.load(sprintf(url, .i),"json")) %>>%
  list.ungroup
```

We have retrieved all the repositories of Hadley Wickham and put them into a list called `repos`. It includes rich information and can be fun to explore with. To know better about the data structure, visit [this page](https://api.github.com/users/hadley/repos) and you will get an impression on how the data is structured.

First, let's see the main languages of the repositories.

```{r}
repos %>>%
  list.table(language) %>>%
  list.sort(desc(.))
```

It is not surprising that most repo's main language is R. Then we can build a much longer pipeline to show the mostly watched active repos. Here we regard the repos updated in the last 30 days as active.

```{r most-watched-repos,fig.height=5,fig.width=12}
repos %>>% 
  list.update(updated_at = updated_at %>>% 
      strptime("%Y-%m-%dT%H:%M:%SZ")) %>>%
  list.filter(updated_at >= as.POSIXct(Sys.Date() - 30)) %>>%
  (~ active_repos) %>>%
  list.sort(desc(watchers)) %>>%
  list.take(10) %>>%
  list.select(name, watchers) %>>%
  list.stack %>>%
  (~ most_watched_repos) %>>%
  with(
    barplot(watchers,names.arg = name,
      main = "Top 10 packages with most watchers (total: %d)" %>>%
        sprintf(sum(watchers)))
    )
```

Then we can use similar way to see repos mainly written in R with most forks. Note that in the previous code, we save the filtered results to `active_repos` for further use so that we don't have to transform the data every time.

```{r most-fork-repos,fig.height=5,fig.width=14}
active_repos %>>%
  list.filter(language == "R") %>>%
  list.sort(desc(forks)) %>>%
  list.take(10) %>>%
  list.select(name, forks) %>>%
  list.stack %>>%
  with(
    barplot(forks,names.arg = name,
      main = "Top 10 packages with most forks (total: %d)" %>>%
        sprintf(sum(forks)))
    )
```

We are also interested in the repos whose name contains *plot* or *vis*. Let's see what they are arranged by popularity measured in count of stargazers.

```{r}
repos %>>% 
  list.filter(!fork) %>>%
  list.filter(equal("plot", name, pattern = TRUE) | 
      equal("vis", name, pattern = TRUE)) %>>%
  list.sort(desc(stargazers_count)) %>>%
  list.mapv(name)
```

If you don't know much about Hadley's repos and heard a repo that sounds like *player*, you can filter the names by fuzzy searching using soundex measurement provided by [stringdist](https://github.com/markvanderloo/stringdist).

```{r}
repos %>>% 
  list.filter(!fork) %>>%
  list.filter(equal("player", name, dist = 0, method = "soundex")) %>>%
  list.mapv(name)
```

Now we know that package is actually spelled as *plyr*.
